---
title: 'Lab #3'
author: "Jay Maxwell"
output: html_document
---
<h4>DS202</h4>
<h4>Fall 2020</h4>
<hr>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
dat <- readxl::read_xls('maxwellGSS.xls')
```

```{r}
##  DATA CLEANING   ##
#  RENAME THE COLUMNS  
# change to factors where needed
# get rid of responses we don't need
# drop all the na
`%notin%` = function(x,y) !(x %in% y)


dat <- dat %>% 
  rename(party='Political party affiliation',
                      edu = 'Highest year of school completed',
                      married = 'Marital status',
                      age = 'Age of respondent',
                      income = 'Total family income',
                      happy = 'General happiness') %>% 
  mutate(party = factor(party), 
         edu=factor(edu), 
         married=factor(married),
         age=factor(age),
         income=factor(income), 
         happy=factor(happy)) %>%
  filter(happy %notin% c("Don't know", "No answer", "Not applicable"),
         married %notin% c('Source','No answer'),
         edu %notin% c("Don't know", 'No answer')) %>%
  drop_na()
```

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.

```{r}
dat1 <- dat %>% 
  group_by(party) %>%
  mutate(edu_avg = mean(as.numeric(edu))) %>% 
  select(party,edu_avg) %>% 
  distinct(party, .keep_all = TRUE ) %>% 
  arrange(desc(edu_avg))

dat1

#dat1
########################################################
# Jaeyeon's Review and Revision
# You can also try this Jay!
# dat$edu = as.factor(dat$edu)
# dat$party = as.factor(dat$party)
# 
# `%notin%` <- Negate(`%in%`)
# 
# jaeyeonDat <- dat %>% 
#   group_by(party) %>%
#   filter(edu %notin% c("No answer", "Don't know")) %>%
#   summarise(mean_edu = mean(as.numeric(edu))) %>%
#   arrange(desc(mean_edu))
# 
# jaeyeonDat
#############################################################
```


2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

```{r}
dat2 <- dat %>%
  group_by(party) %>%
  mutate(edu_med = median(as.numeric(edu))) %>%
  ungroup() %>%
  filter(as.numeric(edu) > edu_med)

str(dat2)

```

3. How is the happiness of a respondent related to his/her marriage status? Define that a person is happy if the response to question is “Very happy” or “Pretty happy”.


```{r}
dat3 <- dat %>% select(happy,married,party)

dat3 <- dat3 %>% group_by(happy,married) %>% summarize(n=n())




q3plot <- ggplot(dat3, aes(x=married, y=n)) + geom_bar(aes(fill=happy),
                                               stat="identity",
                                               color="white",
                                               position = position_dodge(.9)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  xlab("Marital status") + ylab('Frequency') 

q3plot + coord_flip() 

```




4. Does party affiliation affect the relationship you found in the last question?

```{r}
dat3 <- dat %>% 
  select(happy,married,party) %>% 
  group_by(happy,married) %>%
  summarize(n=n()) %>% 
  ungroup()

q3plot <- ggplot(dat3, aes(x=married, y=n)) + 
  geom_bar(aes(fill=happy),
           stat="identity",
           color="white",
           position = position_dodge(.9)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  xlab("Marital status") + ylab('Frequency') 

q3plot #  + facet_grid(~party)
```

I want to do a facet wrap with this, but I get an error. For some reason the "party" data is not available for plotting purposes?


5. How is marrital status related to the education of a respondent? What are possible reasons accounting for the patterns you see?


```{r}

dat5 <- dat %>%
  group_by(married) %>%
  select(married,edu) %>% 
  summarize(avg_edu=mean(as.numeric(edu), n=n())) %>%
  arrange(avg_edu)


dat5 %>% ggplot(aes(x =married, weight = avg_edu)) + geom_bar() + ylab("Average Education in Years") +xlab("Marital Status")

# dat5 %>% ggplot(aes(x = married, weight = n)) +
#   geom_bar()
```

