---
title: 'Lab #3'
author: "Jay Maxwell"
output: html_document
---
<h4>DS202</h4>
<h4>Fall 2020</h4>
<hr>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
dat <- readxl::read_xls('maxwellGSS.xls')
```

```{r}
##  DATA CLEANING   ##
#  RENAME THE COLUMNS  & change to factors where needed
dat <- dat %>% 
  rename(party='Political party affiliation',
                      edu = 'Highest year of school completed',
                      married = 'Marital status',
                      age = 'Age of respondent',
                      income = 'Total family income',
                      happy = 'General happiness') %>% 
  mutate(party = factor(party), 
         edu=factor(edu), 
         married=factor(married),
         age=factor(age),
         income=factor(income), 
         happy=factor(happy))

## TO DO ___
##  GET RID OF 'Don't know'   'No answer'  'NA'   'Not applicable' from happy
## GET RID OF NA    Source     No answer    from married

```

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.

```{r}
#<<<<<<< HEAD
#dat <- dat %>% rename(party='Political party affiliation', edu = 'Highest year of school completed' ) %>% mutate(party = factor(party), edu=factor(edu))
#=======
#>>>>>>> 119dadeb64853e57426df32f519580fc6bc78e27

#dat1 <- dat %>% 
#  group_by(party) %>%
#  mutate(edu_avg = mean(as.numeric(edu))) %>% 
#  select(party,edu_avg) %>% 
#  distinct(party, .keep_all = TRUE ) %>% 
#  arrange(desc(edu_avg))

#dat1

# Jaeyeon's Review and Revision
# You can also try this Jay!
dat$edu = as.factor(dat$edu)
dat$party = as.factor(dat$party)

`%notin%` <- Negate(`%in%`)

jaeyeonDat <- dat %>% 
  group_by(party) %>%
  filter(edu %notin% c("No answer", "Don't know")) %>%
  summarise(mean_edu = mean(as.numeric(edu))) %>%
  arrange(desc(mean_edu))

jaeyeonDat

```


2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

```{r}
dat2 <- dat %>%
  group_by(party) %>%
  mutate(edu_med = median(as.numeric(edu))) %>%
  ungroup() %>%
  filter(as.numeric(edu) > edu_med)

str(dat2)

```

3. How is the happiness of a respondent related to his/her marriage status? Define that a person is happy if the response to question is “Very happy” or “Pretty happy”.


```{r}
#dat3 <- dat %>% select(happy,married,party)

#dat3 <- dat3 %>% group_by(happy,married) %>% summarize(n=n()) 


#dat3

#q3plot <- ggplot(dat3, aes(x=married, y=n)) + geom_bar(aes(fill=happy),
#                                               stat="identity",
#                                               color="white",
#                                               position = position_dodge(.9)) +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
#  xlab("Marital status") + ylab('Frequency') 

#q3plot + coord_flip() facet_wrap(~party)



```




4. Does party affiliation affect the relationship you found in the last question?

```{r}
#<<<<<<< HEAD
#dat4 <- dat %>% 
#  rename(happy = 'General happiness', married='Marital status') %>%
#  mutate(happy = factor(happy), married=factor(married), party=factor(party))

#p <- ggplot(dat4, aes(x=married, y=happy)) + geom_boxplot()
#p

#dat4 <- dat4 %>% 
#  mutate(happy= as.numeric(happy)) %>% 
#  group_by(married,party) %>% 
#  summarize(happy=mean(happy)) %>%
#  arrange(desc(happy))

#dat4  %>% data.frame
  
  
#=======

#>>>>>>> 119dadeb64853e57426df32f519580fc6bc78e27
```

I don't think this works the way I thought it worked. We need to change this. don't sumbit this code!
