---
title: 'Lab #3'
author: "Jay Maxwell"
output: html_document
---
<h4>DS202</h4>
<h4>Fall 2020</h4>
<hr>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
dat <- readxl::read_xls('maxwellGSS.xls')
```

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.

```{r}
#dat <- dat %>% rename(party='Political party affiliation', edu = 'Highest year of school completed' ) %>% mutate(party = factor(party), edu=factor(edu))

#dat1 <- dat %>% 
#  group_by(party) %>%
#  mutate(edu_avg = mean(as.numeric(edu))) %>% 
#  select(party,edu_avg) %>% 
#  distinct(party, .keep_all = TRUE ) %>% 
#  arrange(desc(edu_avg))

#dat1

# Jaeyeon's Review and Revision
# You can also try this Jay!
dat = dat %>% 
  rename(married = 'Marital status',
         age = 'Age of respondent',
         edu = 'Highest year of school completed',
         income = 'Total family income',
         party = 'Political party affiliation', 
         happiness = 'General happiness')

dat$edu = as.factor(dat$edu)
dat$party = as.factor(dat$party)

`%notin%` <- Negate(`%in%`)



jaeyeonDat <- dat %>% 
  group_by(party) %>%
  filter(edu %notin% c("No answer", "Don't know")) %>%
  summarise(mean_edu = mean(as.numeric(edu))) %>%
  arrange(desc(mean_edu))

jaeyeonDat

```


2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

```{r}
dat2 <- dat %>% 
  group_by(party) %>%
  mutate(edu_med = median(as.numeric(edu))) %>%
  ungroup() %>%
  filter(as.numeric(edu) > edu_med)

str(dat2)

```

3. asdfasdfsadf



4. asdfasdfasdf
```{r}
#dat4 <- dat %>% 
#  rename(happy = 'General happiness', married='Marital status') %>%
#  mutate(happy = factor(happy), married=factor(married), party=factor(party))

#p <- ggplot(dat4, aes(x=married, y=happy)) + geom_boxplot()
#p

#dat4 <- dat4 %>% 
#  mutate(happy= as.numeric(happy)) %>% 
#  group_by(married,party) %>% 
#  summarize(happy=mean(happy)) %>%
#  arrange(desc(happy))

#dat4  %>% data.frame
  
  
```
