---
title: 'Lab #3'
author: "Jay Maxwell"
output: html_document
---
<h4>DS202</h4>
<h4>Fall 2020</h4>
<hr>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
dat <- readxl::read_xls('maxwellGSS.xls')
```

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.

```{r}
dat <- dat %>% rename(party='Political party affiliation', edu = 'Highest year of school completed' ) %>% mutate(party = factor(party), edu=factor(edu))

dat1 <- dat %>% 
  group_by(party) %>%
  mutate(edu_avg = mean(as.numeric(edu))) %>% 
  select(party,edu_avg) %>% 
  distinct(party, .keep_all = TRUE ) %>% 
  arrange(desc(edu_avg))

dat1

```


2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

```{r}
dat2 <- dat %>% 
  group_by(party) %>%
  mutate(edu_med = median(as.numeric(edu))) %>%
  ungroup() %>%
  filter(as.numeric(edu) > edu_med)

str(dat2)

```


