---
title: 'Lab #3'
author: "Jay Maxwell, Jaeyeon Won, Alec Meyer"
output: html_document
---
<h4>DS202</h4>
<h4>Fall 2020</h4>
<hr>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
dat <- readxl::read_xls('maxwellGSS.xls')
```

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.

```{r}
dat <- dat %>% rename(party='Political party affiliation', edu = 'Highest year of school completed' ) %>% mutate(party = factor(party), edu=factor(edu))

dat1 <- dat %>% 
  group_by(party) %>%
  mutate(edu_avg = mean(as.numeric(edu))) %>% 
  select(party,edu_avg) %>% 
  distinct(party, .keep_all = TRUE ) %>% 
  arrange(desc(edu_avg))

dat1

```


2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

```{r}
dat2 <- dat %>% 
  group_by(party) %>%
  mutate(edu_med = median(as.numeric(edu))) %>%
  ungroup() %>%
  filter(as.numeric(edu) > edu_med)

str(dat2)

```

3. How is the happiness of a respondent related to his/her marriage status? Define that a person is happy if the response to question is “Very happy” or “Pretty happy”.


```{r}
dat3 <- dat %>% 
  rename(happy = 'General happiness', married='Marital status') %>%
  mutate(happy = factor(happy), married=factor(married)) %>%
  select(happy,married)

dat3 <- dat3 %>% group_by(married,happy) %>% summarize(n=n())


dat3

ggplot(dat3, aes(x=married, y=n)) + geom_bar(aes(fill=happy),
                                               stat="identity",
                                               color="white",
                                               position = position_dodge(.9)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  xlab("Marital status") + ylab('Frequency')



```




4. Does party affiliation affect the relationship you found in the last question?
```{r}
dat4 <- dat %>% 
  rename(happy = 'General happiness', married='Marital status') %>%
  mutate(happy = factor(happy), married=factor(married), party=factor(party))

p <- ggplot(dat4, aes(x=married, y=happy)) + geom_boxplot()
p

dat4 <- dat4 %>% 
  mutate(happy= as.numeric(happy)) %>% 
  group_by(married,party) %>% 
  summarize(happy=mean(happy)) %>%
  arrange(desc(happy))

dat4  %>% data.frame
  
```

I don't think this works the way I thought it worked. We need to change this. don't sumbit this code!
